Why do we Eval?\\[2mm]\Large A large-scale study of Eval usage in R

### 2020-12-16 analysis/usage.Rmd
usage

```{r read, message=TRUE, include=F}
read_fst(eval_calls_path) %>% as_tibble() -> E_raw
read_fst(corpus_path) %>% as_tibble() -> C_raw
read_csv(eval_static_path) %>% semi_join(C_raw, by = "package") -> P
```

@@

C_raw -> C
corpus_size <- nrow(P)

nb_runs <- E %>%
  pull(file) %>%
  n_distinct()
overview_table(
  r("Corpus",corpus_size),
  r("NbRuns", nb_runs)
) 
```

@@

```{r exercised-call-site-per-variant}
static_eval <- P %>% filter(call_fun_name == "eval") %>% nrow()
static_evalq <- P %>% filter(call_fun_name == "evalq") %>% nrow()
static_eval.parent <- P %>% filter(call_fun_name == "eval.parent") %>% nrow()
static_local <- P %>% filter(call_fun_name == "local") %>% nrow()

overview_table(
  r("StaticEval", static_eval),
  r("StaticEvalq", static_evalq),
  r("StaticEvalParent", static_eval.parent),
  r("StaticLocal", static_local),
  r("TriggeredEval", trig_eval),
  r("TriggeredEvalq", trig_evalq),
  r("Triggeredvalparent", trig_eval.parent),
  r("Triggeredlocal", trig_local),
  r("Triggeredevalpercent", ratio(trig_eval, static_eval)),
  r("Triggeredevalqpercent", ratio(trig_evalq, static_evalq)),
  r("Triggeredevalqpercent", ratio(trig_eval.parent, static_eval.parent)),
  r("Localevalqpercent", ratio(trig_local, static_local))
) -> TBL
```

@@

We compute the frequency of evals by package. This shows that some packages
hardly use eval while other use it intensily. The numbers are "for all runs",
this means that a popular package may be triggered by more runs and thus
have higher frequencies. We will later control for this by looking at individual 
call sites.

```{r ev-dist, include=FALSE}
eval_calls_table <- E %>% 
  count(ev_package, wt=duplicates, sort = TRUE)
breaks <- c(0, 10^seq(1,8))
group_tags <- cut(eval_calls_table$n,
                  breaks = breaks,
                  include.lowest = TRUE)
ev_summary <- summary(group_tags)

overview_table(
r("bina", ev_summary[[1]]),
r("binb", ev_summary[[2]]),
r("binc", ev_summary[[3]]),
r("bind", ev_summary[[4]]),
r("bine", ev_summary[[5]]),
r("binf", ev_summary[[6]]),
r("bing", ev_summary[[7]]),
r("binh", ev_summary[[8]])) -> TBL
```

@@

```{r ev-call-precise}

more_1000_calls <- eval_calls_table %>% filter(n > 1000) %>% nrow()

overview_table(
  r("eval calls less hundred", ev_summary[[1]] + ev_summary[[2]]),
  r("eval calls more thousand", more_1000_calls)
)
```

@@

TBL <- overview_table(
  # Sites triggered over the whole corpus:
  r("nb eval call sites", sites),
  # How many packages in the corpus have a single triggered call site:
  r("nb one call site", sites_per_package %>% filter(n == 1)),
  r("at least one call site", nrow(sites_per_package)),
  r("packages no calls to eval", corpus_size - nrow(sites_per_package)),
  r("nb median call sites", median(sites_per_package$n)),
  r("max call sites", max_packages),
)
```

@@

```{r exercised-call-sites, include=F}
static_sites <- nrow(P)

@@

@@

```{r nb-eval-calls}
# run = package?
eval_calls_per_package <- E %>%
  count(ev_package, wt = duplicates) %>%
  left_join(runs_package, by = "ev_package")

>>>>

```{r read, message=TRUE, include=F}
# This test should allow to re-run the whole script without reading the
# data again from disk. It assumes the data doesn't change.
if (exists(eval_calls_path) == FALSE) {
  read_fst(eval_calls_path) %>% as_tibble() -> E_raw
  read_fst(corpus_path) %>% as_tibble() -> C_raw
  read_csv(eval_static_path) %>% semi_join(C_raw, by = "package") -> P
}
```

@@

C_raw -> C
corpus_size <- nrow(C)

nb_runs <- E %>%
  pull(file) %>%
  n_distinct()
overview_table(
  r("Corpus",corpus_size),
  r("Nbruns", nb_runs)
) 
```

@@

@@

We compute the frequency of evals by package. This shows that some packages
hardly use eval while other use it intensly. The numbers are "for all runs",
this means that a popular package may be triggered by more runs and thus
have higher frequencies. We will later control for this by looking at individual 
call sites.

```{r ev-dist, include=FALSE}
# The calls per package
pack_calls <- E %>% 
  count(ev_package, wt=duplicates, sort = TRUE)
# Create bins at every power of ten
bins <- cut(pack_calls$n, breaks = c(0, 10^seq(1,8)), include.lowest = TRUE)
sum <- summary(bins)
# Packages with more than 1K calls
Kcalls <- pack_calls %>% filter(n > 1000) %>% nrow()

# The number of runs per package
runs <- E %>%
  group_by(ev_package) %>%
  summarise(nb_runs = n_distinct(file))

# Number of eval calls per package
ecpp <- pack_calls %>% left_join(runs, by = "ev_package") 
stopifnot( nrow(filter(ecpp, n==0)) == 0)
# The package with the most calls
max_calls <- ecpp %>% filter(n == max(n))

overview_table(
r("bina", sum[[1]]),
r("binb", sum[[2]]),
r("binc", sum[[3]]),
r("bind", sum[[4]]),
r("bine", sum[[5]]),
r("binf", sum[[6]]),
r("bing", sum[[7]]),
r("binh", sum[[8]]),
r("Fewcalls", sum[[1]] + sum[[2]]),
r("Manycalls", Kcalls),
r("Maxcalls", max_calls$n),
r("Maxcallspack", max_calls$ev_package)
) -> TBL
```

@@

@@

TBL <- overview_table(
  # Sites triggered over the whole corpus:
  r("nb eval call sites", sites),
  # How many packages in the corpus have a single triggered call site:
  r("nb one call site", sites_per_package %>% filter(n == 1)),
  r("at least one call site", sites_per_package$n),
  r("packages no calls to eval", corpus_size - nrow(sites_per_package)),
  r("nb median call sites", median(sites_per_package$n)),
  r("max call sites", max_packages),
)
```

@@

```{r exercised-call-sites, include=F}

# TODO :: remove?
static_sites <- nrow(P)

@@

And now, by `eval` variant:

```{r exercised-call-site-per-variant}

static_eval <- P %>% filter(call_fun_name == "eval") %>% nrow()
static_evalq <- P %>% filter(call_fun_name == "evalq") %>% nrow()
static_eval.parent <- P %>% filter(call_fun_name == "eval.parent") %>% nrow()
static_local <- P %>% filter(call_fun_name == "local") %>% nrow()

r("static eval", static_eval)
r("static evalq", static_evalq)
r("static eval parent", static_eval.parent)
r("static local", static_local)

r("triggered eval", trig_eval)
r("triggered evalq", trig_evalq)
r("triggered eval parent", trig_eval.parent)
r("triggered local", trig_local)

r("eval percent call site", ratio(trig_eval, static_eval))
r("evalq percent call site", ratio(trig_evalq, static_evalq))
r("eval parent percent call site", ratio(trig_eval.parent, static_eval.parent))
r("local percent call site", ratio(trig_local, static_local))
```

@@

```{r nb-eval-calls}
# run = package?

### TODO REMOVE
eval_calls_per_package <- E %>%
  count(ev_package, wt = duplicates) %>%
  left_join(runs_package, by = "ev_package")

### 2020-12-16 paper/main.tex
Table variant of evals
usage

Figure~\ref{freq} shows the number of \evals generated by each package
over all runs that invoked functions in that package. Out of the \Corpus
packages in the corpus, \EvalCallsLessHundred packages have a low frequency of \eval calls,
with fewer than 100 of them. On the other hand, \EvalCallsMoreThousand packages call eval
more than 1,000 times, with one package (\MaxEvalCallPackage) calling it \MaxEvalCallsRnd times.

>>>>

Figure~\ref{freq} shows the number of \evals generated by each package
over all runs that invoked functions in that package. Out of the
\Corpus packages in the corpus, \Fewcalls packages have a low
frequency of \eval calls, with fewer than 100 of them. On the other
hand, \Manycalls packages call eval more than 1,000 times, with one
package (\Maxcallspack) calling it \Maxcalls times.
